---
title: "Home Ownership Analysis"
author: Casey Breen
---


Summary: Create causal estimates of association between home ownership and longevity using sibling-based fixed effects models. 


## Library packages 

```{r}
## library packages
library(data.table)
library(tidyverse)
library(ipumsr)
library(socviz)
library(broom)
library(cowplot)
library(fixest)
library(here)
source(here("code/helpers.R"))

## read in sibling mortality data 
dmf_sibs <- fread(here("code/analysis_file/censoc_analysis_sibs.csv")) %>% 
  janitor::clean_names()
```

## Standard ABE models (pooled, pooled adjusted, sib-FE adjusted)

```{r}
## dmf calculate birth order based on sibs 
dmf_analysis_sibs <- dmf_sibs %>% 
  mutate(ownership = relevel(as.factor(ownershp), ref = "20")) %>% 
  group_by(serialp_1920) %>% 
  arrange(byear) %>% 
  mutate(birth_order = row_number()) %>% ungroup()
 
<<<<<<< Updated upstream
model1 <- feols(death_age ~ ownership | byear, weights = dmf_analysis_sibs$weight, data = dmf_analysis_sibs)
model2 <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race  + marst + statefip, weights = dmf_analysis_sibs$weight, data = dmf_analysis_sibs)
model3 <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, weights = dmf_analysis_sibs$weight, data = dmf_analysis_sibs)
=======
## fit models 
model1 <- feols(death_age ~ ownership | byear, data = dmf_analysis_sibs)
model2 <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race  + marst + statefip, data = dmf_analysis_sibs)
model3 <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = dmf_analysis_sibs)
>>>>>>> Stashed changes

## model 1 
results <- tidy(model1) %>% 
  mutate(model = "No Controls") %>% 
  add_row(model = "Controls", estimate = NA, std.error = NA) %>% 
  add_row(model = "Controls + Sibling FE", estimate = NA, std.error = NA) %>% 
  mutate(model = factor(model, levels= c("No Controls", "Controls", "Controls + Sibling FE")))

## model 1 results 
home_ownership_regression_plot1 <- results %>% 
  mutate(e65 = estimate,
         lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error) %>% 
  mutate(method = "OLS") %>% 
  ggplot(aes(x = model, y = e65, ymin = lower, ymax = upper, color = model)) + 
  geom_pointrange(size = 1,  fill = "white", shape = 21) + 
  theme_cowplot() + 
  ylim(-0.2, 0.9) + 
  labs(x = "Model",
       y = "Additional Years of Life") + 
  scale_color_manual(values = cud) + 
  theme(legend.position = "NULL") +
   annotate("text", x = 2.3, y = -.05, label = "Renters", fontface = 'italic') + 
  geom_hline(yintercept = 0, color = "grey", linetype =  "dashed")

## fit model 2 
results <- tidy(model1) %>% mutate(model = "No Controls") %>% 
  bind_rows(tidy(model2) %>% mutate(model = "Controls")) %>% 
 add_row(model = "Controls + Sibling FE", estimate = NA, std.error = NA) %>% 
  mutate(model = factor(model, levels= c("No Controls", "Controls", "Controls + Sibling FE")))

## visualize model 2 
home_ownership_regression_plot2 <- results %>% 
  mutate(e65 = estimate,
         lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error) %>% 
  mutate(method = "OLS") %>% 
  ggplot(aes(x = model, y = e65, ymin = lower, ymax = upper, color = model)) + 
  geom_pointrange(size = 1,  fill = "white", shape = 21) + 
  theme_cowplot() + 
  ylim(-0.2, 0.9) + 
  labs(x = "Model",
       y = "Additional Years of Life") + 
  scale_color_manual(values = cud) + 
  theme(legend.position = "NULL") +
   annotate("text", x = 2.3, y = -.05, label = "Renters", fontface = 'italic') + 
  geom_hline(yintercept = 0, color = "grey", linetype =  "dashed")

## fit model 3 
results <- tidy(model1) %>% mutate(model = "No Controls") %>% 
  bind_rows(tidy(model2) %>% mutate(model = "Controls")) %>% 
  bind_rows(tidy(model3) %>% mutate(model = "Controls + Sibling FE")) %>% 
  mutate(model = factor(model, levels= c("No Controls", "Controls", "Controls + Sibling FE")))

## visualize model 3 
home_ownership_regression_plot3 <- results %>% 
  mutate(e65 = estimate,
         lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error) %>% 
  mutate(method = "OLS") %>% 
  ggplot(aes(x = model, y = e65, ymin = lower, ymax = upper, color = model)) + 
  geom_pointrange(size = 1,  fill = "white", shape = 21) + 
  theme_cowplot() + 
  ylim(-0.2, 0.9) + 
  labs(x = "Model",
       y = "Additional Years of Life") + 
  scale_color_manual(values = cud) + 
  theme(legend.position = "NULL") +
  annotate("text", x = 2.3, y = -.05, label = "Renters", fontface = 'italic') + 
  geom_hline(yintercept = 0, color = "grey", linetype =  "dashed")

## save plots 
ggsave(plot = home_ownership_regression_plot1, "figures/home_ownership_models1.png", width = 6, height = 4)
ggsave(plot = home_ownership_regression_plot2, "figures/home_ownership_models2.png", width = 6, height = 4)
ggsave(plot = home_ownership_regression_plot3, "figures/home_ownership_models3.png", width = 6, height = 4)
```

## Causal ABE models — Conservative Subsample (pooled, pooled adjusted, sib-FE adjusted)

```{r}
## Restrict to conservative subsample for both linkages 
dmf_analysis_sibs_conservative <- dmf_analysis_sibs %>% 
  filter(link_abe_exact_conservative_1940_dmf == 1 & link_abe_exact_conservative_1920_1940 == 1) %>% 
  mutate(ownership = relevel(as.factor(ownershp), ref = "20")) %>% 
  group_by(serialp_1920) %>% 
  filter(n() > 1) 

## Calculate birth order  
dmf_analysis_sibs_conservative <- dmf_analysis_sibs_conservative %>% 
  arrange(byear) %>% 
  mutate(birth_order = row_number()) %>% ungroup()
 
<<<<<<< Updated upstream
model1 <- feols(death_age ~ ownership | byear,  weights = dmf_analysis_sibs_conservative$weight_conservative, data = dmf_analysis_sibs_conservative)
model2 <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race  + marst + statefip, weights = dmf_analysis_sibs_conservative$weight_conservative, data = dmf_analysis_sibs_conservative)
model3 <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, weights = dmf_analysis_sibs_conservative$weight_conservative,  data = dmf_analysis_sibs_conservative)
=======
## Fit models 
model1 <- feols(death_age ~ ownership | byear,  data = dmf_analysis_sibs_conservative)
model2 <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race  + marst + statefip,  data = dmf_analysis_sibs_conservative)
model3 <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order,  data = dmf_analysis_sibs_conservative)
>>>>>>> Stashed changes

## Combine results  
results_conservative <- tidy(model1) %>% mutate(model = "No Controls") %>% 
  bind_rows(tidy(model2) %>% mutate(model = "Controls")) %>% 
  bind_rows(tidy(model3) %>% mutate(model = "Controls + Sibling FE")) %>% 
  mutate(model = factor(model, levels= c("No Controls", "Controls", "Controls + Sibling FE")))

## Visualize results 
home_ownership_regression_conservative_plot <- results_conservative %>% 
  mutate(e65 = estimate,
         lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error) %>% 
  mutate(method = "OLS") %>% 
  ggplot(aes(x = model, y = e65, ymin = lower, ymax = upper, color = model)) + 
  geom_pointrange(size = 1,  fill = "white", shape = 21) + 
  theme_cowplot() + 
  ylim(-0.2, 0.9) + 
  labs(x = "Model",
       y = "Additional Years of Life") + 
  scale_color_manual(values = cud) + 
  theme(legend.position = "NULL") +
   annotate("text", x = 2.3, y = -.05, label = "Renters", fontface = 'italic') + 
  geom_hline(yintercept = 0, color = "grey", linetype =  "dashed")

## Save plot 
ggsave(plot = home_ownership_regression_conservative_plot, "figures/home_ownership_models_conservative.png", width = 6, height = 4)
```

## discordant siblings 

```{r}
## Number of discordant sibships 
dmf_analysis_sibs %>% 
  group_by(serialp_1920) %>% 
  summarize(distinct = n_distinct(ownershp)) %>% 
  count(distinct) %>% 
  summarize(prop = n / sum(n))
```


## Full Regression Models 

```{r}
## Run Regression Models 
dmf_analysis_sibs <- dmf_analysis_sibs %>% 
  mutate(racec = case_when(
    race == 100 ~ "White",
    race == 200 ~ "Black",
    TRUE ~ "Other"
  ))

## fit full regression models 
model1 <- feols(death_age ~ ownership | byear, data = dmf_analysis_sibs)
model2 <- feols(death_age ~ ownership + educ_yrs + racec + urban | byear  + occ +   + marst + statefip, data = dmf_analysis_sibs)
model3 <- feols(death_age ~ ownership + educ_yrs + racec + urban | byear + occ + marst + statefip + serialp_1920 + birth_order, data = dmf_analysis_sibs)

etable(model1, model2, model3, tex = T)
```


## Sensitivity analysis 

Create different, more restrictive subsamples and rerun model 3 on each subgroup. The idea is by making such restrictions (e.g., sibling living in the same county) we are able to control for potential confounding. 

```{r}
## byear restriction 
byear_restrict_df <- dmf_analysis_sibs %>%
   group_by(serialp_1920) %>% 
   mutate(byear_range = max(byear) - min(byear)) %>% 
   filter(byear_range < 6)

model_similar_edu <- tidy(feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = byear_restrict_df)) %>% mutate(model = "Birth Year Restriction")

## county restriction 
county_restriction_df <- dmf_analysis_sibs %>% 
  group_by(serialp_1920) %>% 
  filter(n_distinct(countyicp) == 1) %>% 
  ungroup()

model_county <- tidy(feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = county_restriction_df)) %>% mutate(model = "County Restriction")

## marital status restriction 
marst_restriction_df <- dmf_analysis_sibs %>% 
  group_by(serialp_1920) %>% 
  filter(n_distinct(marst) == 1) %>% 
  ungroup()

marst_similar_model <- tidy(feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = marst_restriction_df)) %>% mutate(model = "Marital Status Restriction")


## education status restriction 
educ_restriction_df <- dmf_analysis_sibs %>% 
  group_by(serialp_1920) %>% 
  mutate(educ_range = max(educ_yrs) - min(educ_yrs)) %>% 
  filter(educ_range < 3) %>% 
  ungroup()

educ_similar_model <- tidy(feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = educ_restriction_df)) %>% mutate(model = "Education Restriction")

## full restrictions 
full_restrictions <- dmf_analysis_sibs %>% 
  group_by(serialp_1920) %>% 
  mutate(byear_range = max(byear) - min(byear)) %>% 
  mutate(educ_range = max(educ_yrs) - min(educ_yrs)) %>% 
  filter(byear_range < 6) %>% 
  filter(educ_range < 3) %>% 
  filter(n_distinct(countyicp) == 1) %>% 
  filter(n_distinct(marst) == 1)

full_restrictions_model <- tidy(feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = full_restrictions)) %>% mutate(model = "All Restrictions ")

## combine estimates into one data frame
results_robustness <- model_similar_edu %>% 
  bind_rows(model_county) %>% 
  bind_rows(marst_similar_model) %>% 
  bind_rows(educ_similar_model) %>% 
  bind_rows(full_restrictions_model)

## sensitivity analysis 
sensitivity_analysis_plots <- results_robustness %>% 
  mutate(e65 = estimate,
         lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error) %>% 
  mutate(method = "OLS") %>% 
  ggplot(aes(y = model, x = e65, xmin = lower, xmax = upper)) + 
  geom_line() + 
  geom_pointrange(size = 0.7, shape = 21, fill = "white") + 
  theme_cowplot() + 
  xlim(-.3, 0.9) + 
  labs(x = "Additional Years of Life",
       y = "") + 
  ggsci::scale_color_lancet() + 
  geom_vline( xintercept = 0.400182, linetype = "dashed") + 
  theme(legend.position = "NULL") + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  annotate("text", x = .05, y = 2.5, label = "Renters", fontface = 'italic', angle='90') + 
  geom_vline(xintercept = 0, color = "grey", linetype =  "dashed") + 
  annotate("text", x = .45, y = 2.5, label = "Full Sample", fontface = 'italic', angle='90') 

## save plot 
ggsave(plot = sensitivity_analysis_plots, "figures/home_ownership_sensitivity_analyses.png", width = 7, height = 5.5)
```


```{r}
results_sensitivity_byear <- list()
  
for (i in 1:5) {
  
## byear restriction 
byear_restrict_df <- dmf_analysis_sibs %>%
    filter(byear >= 1904 + i & byear <= 1904 + 4 + i) %>% 
   group_by(serialp_1920) %>% 
  filter(n() > 1)
  

results_sensitivity_byear[[i]] <- tidy(feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = byear_restrict_df)) %>% mutate(model = "byear restriction") %>% 
  mutate(byear = paste0(1904 + i, ", ", 1905 + 4 + i))
}


bind_rows(results_sensitivity_byear) %>% 
  ggplot(aes(x = byear, y = estimate, ymin = estimate - std.error, ymax = estimate + std.error)) + 
  geom_pointrange() + 
  theme_cowplot() + 
   annotate("text", x = 2.3, y = -.05, label = "Renters", fontface = 'italic') + 
  geom_hline(yintercept = 0, color = "grey", linetype =  "dashed")
```


