---
title: "Home Ownership Analysis"
author: Casey Breen
---


Summary: Create causal estimates of association between home ownership and longevity using sibling-based fixed effects models. 


## Library packages 

```{r}
## library packages
library(data.table)
library(tidyverse)
library(ipumsr)
library(socviz)
library(broom)
library(cowplot)
library(fixest)
library(here)
```


## Read in data 

```{r}
## read in sibling mortality data 
dmf_sibs <- fread(here("code/analysis_file/censoc_analysis_sibs.csv")) %>% 
  janitor::clean_names()
```

## Standard ABE models (pooled, pooled adjusted, sib-FE adjusted)

```{r}
dmf_analysis_sibs <- dmf_sibs %>% 
  mutate(ownership = relevel(as.factor(ownershp), ref = "20")) %>% 
  filter(ownership != 0) %>% 
  filter(relate == 101) %>% 
  group_by(serialp_1920) %>% 
  filter(n() > 1) 

dmf_analysis_sibs <- dmf_analysis_sibs %>% 
  group_by(serialp_1920) %>% 
  arrange(byear) %>% 
  mutate(birth_order = row_number()) %>% ungroup()
 
model1 <- feols(death_age ~ ownership | byear, data = dmf_analysis_sibs)
model2 <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race  + marst + statefip, data = dmf_analysis_sibs)
model3 <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = dmf_analysis_sibs)

## model 1 
results <- tidy(model1) %>% mutate(model = "No Controls") %>% 
 add_row(model = "Controls", estimate = NA, std.error = NA) %>% 
 add_row(model = "Controls + Sibling FE", estimate = NA, std.error = NA) %>% 
  mutate(model = factor(model, levels= c("No Controls", "Controls", "Controls + Sibling FE")))

home_ownership_regression_plot1 <- results %>% 
  mutate(e65 = estimate,
         lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error) %>% 
  mutate(method = "OLS") %>% 
  ggplot(aes(x = model, y = e65, ymin = lower, ymax = upper, color = model)) + 
  geom_line() + 
  geom_pointrange(size = 1) + 
  theme_cowplot() + 
  ylim(-0.2, 0.9) + 
  labs(x = "Model",
       y = "Additional Years of Life") + 
  ggsci::scale_color_aaas() + 
  theme(legend.position = "NULL") +
   annotate("text", x = 2.3, y = -.05, label = "Renters", fontface = 'italic') + 
  geom_hline(yintercept = 0, color = "grey", linetype =  "dashed")

## model 2 
results <- tidy(model1) %>% mutate(model = "No Controls") %>% 
  bind_rows(tidy(model2) %>% mutate(model = "Controls")) %>% 
 add_row(model = "Controls + Sibling FE", estimate = NA, std.error = NA) %>% 
  mutate(model = factor(model, levels= c("No Controls", "Controls", "Controls + Sibling FE")))

home_ownership_regression_plot2 <- results %>% 
  mutate(e65 = estimate,
         lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error) %>% 
  mutate(method = "OLS") %>% 
  ggplot(aes(x = model, y = e65, ymin = lower, ymax = upper, color = model)) + 
  geom_line() + 
  geom_pointrange(size = 1) + 
  theme_cowplot() + 
  ylim(-0.2, 0.9) + 
  labs(x = "Model",
       y = "Additional Years of Life") + 
  ggsci::scale_color_aaas() + 
  theme(legend.position = "NULL") +
   annotate("text", x = 2.3, y = -.05, label = "Renters", fontface = 'italic') + 
  geom_hline(yintercept = 0, color = "grey", linetype =  "dashed")

## model 3 
results <- tidy(model1) %>% mutate(model = "No Controls") %>% 
  bind_rows(tidy(model2) %>% mutate(model = "Controls")) %>% 
  bind_rows(tidy(model3) %>% mutate(model = "Controls + Sibling FE")) %>% 
  mutate(model = factor(model, levels= c("No Controls", "Controls", "Controls + Sibling FE")))

home_ownership_regression_plot3 <- results %>% 
  mutate(e65 = estimate,
         lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error) %>% 
  mutate(method = "OLS") %>% 
  ggplot(aes(x = model, y = e65, ymin = lower, ymax = upper, color = model)) + 
  geom_line() + 
  geom_pointrange(size = 1) + 
  theme_cowplot() + 
  ylim(-0.2, 0.9) + 
  labs(x = "Model",
       y = "Additional Years of Life") + 
  ggsci::scale_color_aaas() + 
  theme(legend.position = "NULL") +
   annotate("text", x = 2.3, y = -.05, label = "Renters", fontface = 'italic') + 
  geom_hline(yintercept = 0, color = "grey", linetype =  "dashed")

ggsave(plot = home_ownership_regression_plot1, "figures/home_ownership_models1.png", width = 6, height = 4)
ggsave(plot = home_ownership_regression_plot2, "figures/home_ownership_models2.png", width = 6, height = 4)
ggsave(plot = home_ownership_regression_plot3, "figures/home_ownership_models3.png", width = 6, height = 4)
```

## Causal ABE models (pooled, pooled adjusted, sib-FE adjusted)

```{r}
dmf_analysis_sibs_conservative <- dmf_analysis_sibs %>% 
  filter(link_abe_exact_conservative_1940_dmf == 1 & link_abe_exact_conservative_1920_1940 == 1) %>% 
  mutate(ownership = relevel(as.factor(ownershp), ref = "20")) %>% 
  filter(ownership != 0) %>% 
  filter(relate == 101) %>% 
  group_by(serialp_1920) %>% 
  filter(n() > 1) 

dmf_analysis_sibs_conservative <- dmf_analysis_sibs_conservative %>% 
  arrange(byear) %>% 
  mutate(birth_order = row_number()) %>% ungroup()
 
model1 <- feols(death_age ~ ownership | byear,  data = dmf_analysis_sibs_conservative)
model2 <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race  + marst + statefip,  data = dmf_analysis_sibs_conservative)
model3 <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order,  data = dmf_analysis_sibs_conservative)

results_conservative <- tidy(model1) %>% mutate(model = "No Controls") %>% 
  bind_rows(tidy(model2) %>% mutate(model = "Controls")) %>% 
  bind_rows(tidy(model3) %>% mutate(model = "Controls + Sibling FE")) %>% 
  mutate(model = factor(model, levels= c("No Controls", "Controls", "Controls + Sibling FE")))

home_ownership_regression_conservative_plot <- results_conservative %>% 
  mutate(e65 = estimate,
         lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error) %>% 
  mutate(method = "OLS") %>% 
  ggplot(aes(x = model, y = e65, ymin = lower, ymax = upper, color = model)) + 
  geom_line() + 
  geom_pointrange(size = 1) + 
  theme_cowplot() + 
  ylim(-0.2, 0.9) + 
  labs(x = "Model",
       y = "Additional Years of Life") + 
  ggsci::scale_color_aaas() + 
  theme(legend.position = "NULL") +
   annotate("text", x = 2.3, y = -.05, label = "Renters", fontface = 'italic') + 
  geom_hline(yintercept = 0, color = "grey", linetype =  "dashed")

# title = "Relationship between Home Ownership and Longevity",
# subtitle = "CenSoc-DMF cohorts of 1905-1915, Men-Only"


ggsave(plot = home_ownership_regression_conservative_plot, "figures/home_ownership_models_conservative.pdf", width = 6, height = 4)
```

## Value of home analysis (sib FE)

```{r}
dmf_analysis_sibs_new <- dmf_analysis_sibs %>% 
  filter(ownership != 10 | valueh != 9999998) %>% 
  mutate(ownership_value = case_when(
    ownership == 10 & valueh %in% c(1:700) ~ 1,
    ownership == 10 & valueh %in% c(700:2000) ~ 2,
    ownership == 10 & valueh %in% c(2000:3500) ~ 3,
    ownership == 10 & valueh %in% c(3500:10000) ~ 4,
    TRUE ~ 0
    ))

## run models 
model1 <- tidy(feols(death_age ~ as.factor(ownership_value)  | byear,  data = dmf_analysis_sibs_new)) %>% mutate(model = "No Controls")
model2 <- tidy(feols(death_age ~ as.factor(ownership_value)  | byear + educ_yrs + urban + occ + race + marst + statefip,  data = dmf_analysis_sibs_new)) %>% mutate(model = "Controls")
model3 <- tidy(feols(death_age ~ as.factor(ownership_value)  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920,  data = dmf_analysis_sibs_new)) %>% mutate(model = "Controls + Sibling FE")

## effect heterogeneity 
home_ownership_homevalue_plot_1models <- model3 %>% 
  # bind_rows(model2) %>% 
  # bind_rows(model1) %>% 
  mutate(term = case_when(
    term == "as.factor(ownership_value)1" ~ "Q1 (Low)",
    term == "as.factor(ownership_value)2" ~ "Q2",
    term == "as.factor(ownership_value)3" ~ "Q3",
    term == "as.factor(ownership_value)4" ~ "Q4 (High)"
  )) %>% 
  mutate(e65 = estimate,
         lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error) %>% 
  ggplot(aes(x = term, y = e65, ymin = lower, ymax = upper, group = model, color = model, shape = model)) + 
  geom_pointrange(size = 0.7, position = position_dodge2(.1)) + 
  geom_line(alpha = 0.6, linetype = "dashed") + 
  theme_cowplot() + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey") + 
  labs(x = "Home Value Quartile",
       y = "Additional Years of Life") + 
 #  ggsci::scale_color_lancet() + 
  scale_color_manual(values = c("Controls + Sibling FE" = "#008B45FF")) + 
  # scale_color_viridis_d(option = "m") + 
  theme(legend.position = "bottom") + 
    annotate("text", x = 3.5, y = -.05, label = "Renters",
  parse = TRUE)

ggsave(plot = home_ownership_homevalue_plot_1models, "figures/home_ownership_homevalue_plot_1models.png", width = 6, height = 4)

## effect heterogeneity 
home_ownership_homevalue_plot_3models <- model1 %>% 
  bind_rows(model2) %>% 
  bind_rows(model3) %>% 
  mutate(term = case_when(
    term == "as.factor(ownership_value)1" ~ "Q1 (Low)",
    term == "as.factor(ownership_value)2" ~ "Q2",
    term == "as.factor(ownership_value)3" ~ "Q3",
    term == "as.factor(ownership_value)4" ~ "Q4 (High)"
  )) %>% 
  mutate(model = factor(model, levels= c("No Controls", "Controls", "Controls + Sibling FE"))) %>% 
  mutate(e65 = estimate,
         lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error) %>% 
  ggplot(aes(x = term, y = e65, ymin = lower, ymax = upper, group = model, color = model, shape = model)) + 
  geom_pointrange(size = 0.7, position = position_dodge2(.1)) + 
  geom_line(alpha = 0.6, linetype = "dashed") + 
  theme_cowplot() + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey") + 
    ggsci::scale_color_aaas() + 
  labs(x = "Home Value Quartile",
       y = "Additional Years of Life") + 
  # scale_color_viridis_d(option = "m") + 
  theme(legend.position = "bottom") + 
    annotate("text", x = 3.5, y = -.05, label = "Renters",
  parse = TRUE)

ggsave(plot = home_ownership_homevalue_plot_3models, "figures/home_ownership_homevalue_plot_3models.png", width = 6, height = 4)
```

## discordant siblings 

```{r}
## Number of discordant sibships 
dmf_analysis_sibs %>% 
  group_by(serialp_1920) %>% 
  summarize(distinct = n_distinct(ownershp)) %>% 
  count(distinct) %>% 
  summarize(prop = n / sum(n))
```


## Full Rgression Models 

```{r}
dmf_analysis_sibs <- dmf_analysis_sibs %>% 
  mutate(racec = case_when(
    race == 100 ~ "White",
    race == 200 ~ "Black",
    TRUE ~ "Other"
  ))

model1 <- feols(death_age ~ ownership | byear, data = dmf_analysis_sibs)
model2 <- feols(death_age ~ ownership + educ_yrs + racec + urban | byear  + occ +   + marst + statefip, data = dmf_analysis_sibs)
model3 <- feols(death_age ~ ownership + educ_yrs + racec + urban | byear + occ + marst + statefip + serialp_1920 + birth_order, data = dmf_analysis_sibs)

etable(model1, model2, model3, tex = T)
```


## Sensitivity analysis 

```{r}
subgroup <- dmf_analysis_sibs %>% 
  group_by(serialp_1920) %>% 
  filter(n_distinct(countyicp) == 1) 

model3 <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = subgroup)

subgroup_lowedu <- dmf_analysis_sibs %>% 
  filter(educ_yrs < 9) %>% 
  group_by(serialp_1920) %>% 
  filter(n() >= 1)

subgroup_highedu <- dmf_analysis_sibs %>% 
  filter(educ_yrs >= 9) %>% 
  group_by(serialp_1920) %>% 
  filter(n() >= 1)

model_highedu <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = subgroup_highedu)

model_lowedu <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = subgroup_lowedu)


subgroup <- dmf_analysis_sibs %>%
  filter(!is.na(educ_yrs)) %>% 
  group_by(serialp_1920) %>% 
  filter(n_distinct(countyicp) == 1) %>% 
  group_by(serialp_1920) %>% 
  mutate(educ_range = max(educ_yrs) - min(educ_yrs)) %>% 
  mutate(byear_range = max(byear) - min(byear)) %>% 
  filter(educ_range < 3) %>% 
  filter(byear_range < 3)

model_similar_edu <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = subgroup)

model_similar_edu <- feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = subgroup)


## byear restriction 
byear_restrict_df <- dmf_analysis_sibs %>%
   group_by(serialp_1920) %>% 
   mutate(byear_range = max(byear) - min(byear)) %>% 
   filter(byear_range < 6)

model_similar_edu <- tidy(feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = byear_restrict_df)) %>% mutate(model = "Birth Year Restriction")

## county restriction 
county_restriction_df <- dmf_analysis_sibs %>% 
  group_by(serialp_1920) %>% 
  filter(n_distinct(countyicp) == 1) %>% 
  ungroup()

model_county <- tidy(feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = county_restriction_df)) %>% mutate(model = "County Restriction")

## marital status restriction 
marst_restriction_df <- dmf_analysis_sibs %>% 
  group_by(serialp_1920) %>% 
  filter(n_distinct(marst) == 1) %>% 
  ungroup()

marst_similar_model <- tidy(feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = marst_restriction_df)) %>% mutate(model = "Marital Status Restriction")


## education status restriction 
educ_restriction_df <- dmf_analysis_sibs %>% 
  group_by(serialp_1920) %>% 
  mutate(educ_range = max(educ_yrs) - min(educ_yrs)) %>% 
  filter(educ_range < 3) %>% 
  ungroup()

educ_similar_model <- tidy(feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = educ_restriction_df)) %>% mutate(model = "Education Restriction")


full_restrictions <- dmf_analysis_sibs %>% 
  group_by(serialp_1920) %>% 
  mutate(byear_range = max(byear) - min(byear)) %>% 
  mutate(educ_range = max(educ_yrs) - min(educ_yrs)) %>% 
  filter(byear_range < 6) %>% 
  filter(educ_range < 3) %>% 
  filter(n_distinct(countyicp) == 1) %>% 
  filter(n_distinct(marst) == 1)

full_restrictions_model <- tidy(feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = full_restrictions)) %>% mutate(model = "All Restrictions ")

results_robustness <- model_similar_edu %>% 
  bind_rows(model_county) %>% 
  bind_rows(marst_similar_model) %>% 
  bind_rows(educ_similar_model) %>% 
  bind_rows(full_restrictions_model)

sensitivity_analysis_plots <- results_robustness %>% 
  mutate(e65 = estimate,
         lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error) %>% 
  mutate(method = "OLS") %>% 
  ggplot(aes(y = model, x = e65, xmin = lower, xmax = upper)) + 
  geom_line() + 
  geom_pointrange(size = 0.7) + 
  theme_cowplot() + 
  xlim(-.3, 0.9) + 
  labs(x = "Additional Years of Life",
       y = "") + 
  ggsci::scale_color_lancet() + 
  geom_vline( xintercept = 0.400182, linetype = "dashed") + 
  theme(legend.position = "NULL") + 
  annotate("text", x = .05, y = 2.5, label = "Renters", fontface = 'italic', angle='90') + 
  geom_vline(xintercept = 0, color = "grey", linetype =  "dashed") + 
  annotate("text", x = .45, y = 2.5, label = "Full Sample", fontface = 'italic', angle='90') 


# title = "Relationship between Home Ownership and Longevity",
# subtitle = "CenSoc-DMF cohorts of 1905-1915, Men-Only"

ggsave(plot = sensitivity_analysis_plots, "figures/home_ownership_sensitivity_analyses.png", width = 7, height = 5.5)

```


```{r}
results_sensitivity_byear <- list()
  
for (i in 1:5) {
  
## byear restriction 
byear_restrict_df <- dmf_analysis_sibs %>%
    filter(byear >= 1904 + i & byear <= 1904 + 4 + i) %>% 
   group_by(serialp_1920) %>% 
  filter(n() > 1)
  

results_sensitivity_byear[[i]] <- tidy(feols(death_age ~ ownership  | byear + educ_yrs + urban + occ + race + marst + statefip + serialp_1920 + birth_order, data = byear_restrict_df)) %>% mutate(model = "byear restriction") %>% 
  mutate(byear = paste0(1904 + i, ", ", 1905 + 4 + i))
}


bind_rows(results_sensitivity_byear) %>% 
  ggplot(aes(x = byear, y = estimate, ymin = estimate - std.error, ymax = estimate + std.error)) + 
  geom_pointrange() + 
  theme_cowplot() + 
   annotate("text", x = 2.3, y = -.05, label = "Renters", fontface = 'italic') + 
  geom_hline(yintercept = 0, color = "grey", linetype =  "dashed")
```


